Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2018-03-28 revised_prop_letter_ajhs.tex
revised_prop_letter_ajhs.tex \newenvironment{oldText}[1]{\begin{compareText}{#1}{old}}{\end{compareText}\newpage}
revised_prop_letter_ajhs.tex The version before meeting with Andy...

% https://tex.stackexchange.com/questions/303018/minipage-reset-indent-how-can-i-set-again-to-standard
\edef\docIndent{\the\parindent}
\edef\docParskip{\the\parskip}
\newenvironment{docParinfo}{%
 \edef\currentIndent{\the\parindent}
 \edef\currentParskip{\the\parskip}
 \setlength{\parindent}{\docIndent}%
 \setlength{\parskip}{\docParskip}
}{
 \setlength{\parindent}{\currentIndent}%
 \setlength{\parskip}{\currentParskip}
}
\newcommand{\says}[3]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{#1}{#2 says: #3}\end{minipage}\end{framed}}
\newcommand{\amr}[1]{\says{green}{Amr}{#1}}
\newcommand{\yutsung}[1]{\says{purple}{Yu-Tsung}{#1}}
\newcommand{\gerardo}[1]{\says{OliveGreen}{Gerardo}{#1}}
\newcommand{\andy}[1]{\says{blue}{Andy}{#1}}
% \excludecomment{suggest}
\NewDocumentEnvironment{suggest}{m m O{}}{#1{TEXT \##2 would go here. #3}}{#1{TEXT \##2 end here.}}
\newenvironment{compare}{\begin{widetext}\end{widetext}}{\begin{widetext}\end{widetext}}
\newenvironment{compareNextpage}{\clearpage}{\begin{widetext}\end{widetext}}
\NewDocumentEnvironment{compareText}{m m}{\noindent\begin{minipage}[b]{1\columnwidth}#1{The following is #2 text.}\begin{docParinfo}}{\end{docParinfo}#1{The #2 text end here.}\end{minipage}}
\newenvironment{oldText}[1]{\begin{compareText}{#1}{old}}{\end{compareText}\newpage}
\newenvironment{newText}[1]{\begin{compareText}{#1}{new}}{\end{compareText}}

@@

\begin{compare}
\begin{oldText}{\andy}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  {\color{blue}Note that the possibility of negative
intervals appearing in Eq.~(\ref{eq:interval-times})
is a natural consequence of the minus sign appearing in
Eq.~(\ref{eq:complement}).}  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
aspects of both classical IVPMs and quantum probability measures.
\end{oldText}
\begin{newText}{\yutsung}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  {\color{purple}Note that we need to
consider $x\le0$ case in Eq.~(\ref{eq:interval-times}) because of the minus sign
appearing in
Eq.~(\ref{eq:complement}).}  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
aspects of both classical IVPMs and quantum probability measures.
\end{newText}
\end{compare}

Our definition of QIVPMs is consistent with classical IVPMs in the
sense that a restriction of QIVPMs to mutually commuting events,
$\eventsC$, recovers the definition of classical
IVPMs~\cite{JamisonLodwick2004}. The proof of this fact is included in the
forthcoming thesis by the first author~\cite{TaiThesis2018}. A
consequence is that known properties of classical IVPMs
directly hold for QIVPMs when one restricts to mutually commuting
events, $\eventsC$. In particular, in the classical world, it is
\emph{impossible} for experiments to result in probabilities that are
inconsistent with \emph{some} state of the system under consideration,
i.e., all IVPMs must have a non-empty ``core''~\footnote{A result by
  Shapley~\cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997,Grabisch2016}
  proves that a classical IVPM always contains at least one state that
  is consistent with every event.}. Interestingly, as we show in
Sec.~\ref{sec:Gleason}, it is possible in the quantum world for the
probabilities associated with some events to be inconsistent with
\emph{any} quantum state, i.e., for the QIVPM to have an empty core;
in that case, one cannot guarantee non-empty cores for finite-precision
attempts at proving Gleason's theorem by extending the Born measure
$\muB_{\rho}\left(P\right)$ to QIVPMs
$\bar{\mu}\left(P\right)$. However, if we restrict ourselves to the set
$\eventsC$ of mutually
commuting, the situation reverts to the classical case in which
probabilities always determine at least one state.

@@

We next refine the concept of consistency by introducing the idea of
a ``core'' set of states relative to subspaces. First, we define
$\events'$ as a \emph{subspace} of a set of events if $\events$ if
$\events'$ contains the projector $\mathbb{0}$ and $\mathbb{1}$ and is closed under
complements, sums, and products.  In particular, for any
projector~$P\in\events'$, we have
$\mathbb{1}-P\in\events'$ and for each pair of commuting
projectors~$P_{0}\in\events'$ and $P_{1}\in\events'$, we have
$P_{0}+P_{1}-P_{0}P_{1}$ and $P_{0}P_{1}\in\events'$.

@@

\begin{compare}
\begin{oldText}{\andy}
To that end, we construct a family of QIVPMs parameterized by {\color{blue}\emph{blurring}
functions. A blurring function} $f:\left[0,1\right]\rightarrow\mathscr{I}$
maps every real-valued probability $x\in\left[0,1\right]$ to a set
of intervals $f\left(x\right)=\left[\ell,r\right]$ containing $x$,
where $\left[0,1\right]$ denotes the set of real-valued probabilities
which should not be considered as the interval-valued probability
$\unknown$. We also need a notion of norm to quantify the
distance between (pure or mixed) states. The norm of a pure state
$\rho=\proj{\psi}$ is defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
In addition, given any Hermitian operator~$A$, we consider the usual
operator norm $\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
the norm is the maximum absolute value of its eigenvalue. Then, a
finite-precision extension of Gleason's theorem can be stated as follows.
\end{oldText}
\begin{newText}{\andy}
To that end, we construct a family of QIVPMs parameterized by {\color{blue}\emph{interval}
maps. A interval map} $f:\left[0,1\right]\rightarrow\mathscr{I}$
maps every real-valued probability $x\in\left[0,1\right]$ to a set
of intervals $f\left(x\right)=\left[\ell,r\right]$ containing $x$,
where $\left[0,1\right]$ denotes the set of real-valued probabilities
which should not be considered as the interval-valued probability
$\unknown$. We also need a notion of norm to quantify the
distance between (pure or mixed) states. The norm of a pure state
$\rho=\proj{\psi}$ is defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
In addition, given any Hermitian operator~$A$, we consider the usual
operator norm $\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
the norm is the maximum absolute value of its eigenvalue. Then, a
finite-precision extension of Gleason's theorem can be stated as follows.
\end{newText}
\end{compare}

\begin{compareNextpage}
\begin{oldText}{\andy}
\begin{thm}[Finite-Precision Extension of the Gleason Theorem]\label{thm:Finite-precision-Gleason}Let
$f:\left[0,1\right]\rightarrow\mathscr{I}$ be a {\color{blue}blurring function}
and let $f\circ\muB_{\rho}$ be a QIVPM, where $\muB_{\rho}$ is the
probability measure induced by the Born rule for a given state~$\rho$.
Let $\alpha$ be the maximum length of intervals in $\mathscr{I}$.
If a state $\rho'$ is consistent with $f\circ\muB_{\rho}$ on all
events, i.e., $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
then the norm of their difference is bounded by $\alpha$, i.e., $\left\Vert \rho-\rho'\right\Vert \le\alpha$.\end{thm}
\end{oldText}
\begin{newText}{\andy}
\begin{thm}[Finite-Precision Extension of the Gleason Theorem]\label{thm:Finite-precision-Gleason}Let
$f:\left[0,1\right]\rightarrow\mathscr{I}$ be a {\color{blue}interval map}
and let $f\circ\muB_{\rho}$ be a QIVPM, where $\muB_{\rho}$ is the
probability measure induced by the Born rule for a given state~$\rho$.
Let $\alpha$ be the maximum length of intervals in $\mathscr{I}$.
If a state $\rho'$ is consistent with $f\circ\muB_{\rho}$ on all
events, i.e., $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
then the norm of their difference is bounded by $\alpha$, i.e., $\left\Vert \rho-\rho'\right\Vert \le\alpha$.\end{thm}
\end{newText}
\end{compareNextpage}

@@

\begin{compare}
\begin{oldText}{\andy}
To summarize, a conventional quantum probability measure has an
uncountable range $[0,1]$. A QIVPM constructed by {\color{blue}blurring} such a
conventional quantum probability measure must also have an uncountable
range of intervals. Of course, any particular QIVPM, or any particular
experiment, will use a fixed collection of intervals appropriate for
the resources and precision of the particular experiment. 
\end{oldText}
\begin{newText}{\andy}
To summarize, a conventional quantum probability measure has an
uncountable range $[0,1]$. A QIVPM constructed by {\color{blue}?????} such a
conventional quantum probability measure must also have an uncountable
range of intervals. Of course, any particular QIVPM, or any particular
experiment, will use a fixed collection of intervals appropriate for
the resources and precision of the particular experiment. 
\end{newText}
\end{compare}

>>>>

% https://tex.stackexchange.com/questions/303018/minipage-reset-indent-how-can-i-set-again-to-standard
\edef\docIndent{\the\parindent}
\edef\docParskip{\the\parskip}
\newenvironment{docParinfo}{%
 \edef\currentIndent{\the\parindent}
 \edef\currentParskip{\the\parskip}
 \setlength{\parindent}{\docIndent}%
 \setlength{\parskip}{\docParskip}
}{
 \setlength{\parindent}{\currentIndent}%
 \setlength{\parskip}{\currentParskip}
}
\newcommand{\says}[3]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{#1}{#2 says: #3}\end{minipage}\end{framed}}
\newcommand{\amr}[1]{\says{green}{Amr}{#1}}
\newcommand{\yutsung}[1]{\says{purple}{Yu-Tsung}{#1}}
\newcommand{\gerardo}[1]{\says{OliveGreen}{Gerardo}{#1}}
\newcommand{\andy}[1]{\says{blue}{Andy}{#1}}
% \excludecomment{suggest}
\NewDocumentEnvironment{suggest}{m m O{}}{#1{TEXT \##2 would go here. #3}}{#1{TEXT \##2 end here.}}
\newenvironment{compare}{\begin{widetext}\end{widetext}}{\begin{widetext}\end{widetext}}
\newenvironment{compareNextpage}{\clearpage}{\begin{widetext}\end{widetext}}
\NewDocumentEnvironment{compareText}{m m}{\noindent\begin{minipage}[b]{1\columnwidth}#1{The following is #2 text.}\begin{docParinfo}}{\end{docParinfo}#1{The #2 text end here.}\end{minipage}}
\newenvironment{oldText}[1]{\begin{compareText}{#1}{old}}{\end{compareText}\penalty-10000}
\newenvironment{newText}[1]{\begin{compareText}{#1}{new}}{\end{compareText}}

@@

\begin{compare}
\begin{oldText}{\andy}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  {\color{blue}Note that the possibility of negative
intervals appearing in Eq.~(\ref{eq:interval-times})
is a natural consequence of the minus sign appearing in
Eq.~(\ref{eq:complement}).}  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
aspects of both classical IVPMs and quantum probability measures.
\end{oldText}
\begin{newText}{\yutsung}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  {\color{purple}Note that we need to
consider $x\le0$ case in Eq.~(\ref{eq:interval-times}) because the minus sign
appears in
Eq.~(\ref{eq:complement}).}  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
aspects of both classical IVPMs and quantum probability measures.
\end{newText}
\end{compare}

Our definition of QIVPMs is consistent with classical IVPMs in the
sense that a restriction of QIVPMs to mutually commuting events,
$\eventsC$, recovers the definition of classical
IVPMs~\cite{JamisonLodwick2004}. The proof of this fact is included in the
forthcoming thesis by the first author~\cite{TaiThesis2018}. A
consequence is that known properties of classical IVPMs
directly hold for QIVPMs when one restricts to mutually commuting
events, $\eventsC$. In particular, in the classical world, it is
\emph{impossible} for experiments to result in probabilities that are
inconsistent with \emph{some} state of the system under consideration,
i.e., all IVPMs must have a non-empty ``core''~\footnote{A result by
  Shapley~\cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997,Grabisch2016}
  proves that a classical IVPM always contains at least one state that
  is consistent with every event.}. Interestingly, as we show in
Sec.~\ref{sec:Gleason}, it is possible in the quantum world for the
probabilities associated with some events to be inconsistent with
\emph{any} quantum state, i.e., for the QIVPM to have an empty core;
in that case, one cannot guarantee non-empty cores for finite-precision
attempts at proving Gleason's theorem by extending the Born measure
$\muB_{\rho}\left(P\right)$ to QIVPMs
$\bar{\mu}\left(P\right)$. However, if we restrict ourselves to the set
$\eventsC$ of mutually commuting, \yutsung{Since ``mutually commuting'' is the
object of the proposition ``of'', is ``mutually commuting'' a noun?}
the situation reverts to the classical case in which
probabilities always determine at least one state.

@@

\begin{compare}
\begin{oldText}{\yutsung}
We next refine the concept of consistency by introducing the idea of
a ``core'' set of states relative to subspaces. First, we define
$\events'$ as a \emph{subspace} of a set of events {\color{purple}if} $\events$ if
$\events'$ contains the projector $\mathbb{0}$ and $\mathbb{1}$ and is closed under
complements, sums, and products.  In particular, for any
projector~$P\in\events'$, we have
$\mathbb{1}-P\in\events'$ and for each pair of commuting
projectors~$P_{0}\in\events'$ and $P_{1}\in\events'$, we have
$P_{0}+P_{1}-P_{0}P_{1}$ and $P_{0}P_{1}\in\events'$.
\end{oldText}
\begin{newText}{\yutsung}
We next refine the concept of consistency by introducing the idea of
a ``core'' set of states relative to subspaces. First, we define
$\events'$ as a \emph{subspace} of a set of events~$\events$ if
$\events'$ contains the projector $\mathbb{0}$ and $\mathbb{1}$ and is closed under
complements, sums, and products.  In particular, for any
projector~$P\in\events'$, we have
$\mathbb{1}-P\in\events'$ and for each pair of commuting
projectors~$P_{0}\in\events'$ and $P_{1}\in\events'$, we have
$P_{0}+P_{1}-P_{0}P_{1}$ and $P_{0}P_{1}\in\events'$.
\end{newText}
\end{compare}

@@

\begin{compare}
\begin{oldText}{\andy}
To that end, we construct a family of QIVPMs parameterized by {\color{blue}\emph{blurring}
functions. A blurring function} $f:\left[0,1\right]\rightarrow\mathscr{I}$
maps every real-valued probability $x\in\left[0,1\right]$ to a set
of intervals $f\left(x\right)=\left[\ell,r\right]$ containing $x$,
where $\left[0,1\right]$ denotes the set of real-valued probabilities
which should not be considered as the interval-valued probability
$\unknown$. We also need a notion of norm to quantify the
distance between (pure or mixed) states. The norm of a pure state
$\rho=\proj{\psi}$ is defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
In addition, given any Hermitian operator~$A$, we consider the usual
operator norm $\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
the norm is the maximum absolute value of its eigenvalue. Then, a
finite-precision extension of Gleason's theorem can be stated as follows.

\begin{thm}[Finite-Precision Extension of the Gleason Theorem]\label{thm:Finite-precision-Gleason}Let
$f:\left[0,1\right]\rightarrow\mathscr{I}$ be a {\color{blue}blurring function}
and let $f\circ\muB_{\rho}$ be a QIVPM, where $\muB_{\rho}$ is the
probability measure induced by the Born rule for a given state~$\rho$.
Let $\alpha$ be the maximum length of intervals in $\mathscr{I}$.
If a state $\rho'$ is consistent with $f\circ\muB_{\rho}$ on all
events, i.e., $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
then the norm of their difference is bounded by $\alpha$, i.e., $\left\Vert \rho-\rho'\right\Vert \le\alpha$.\end{thm}
\end{oldText}
\begin{newText}{\andy}
To that end, we construct a family of QIVPMs parameterized by {\color{blue}\emph{interval}
maps. A interval map} $f:\left[0,1\right]\rightarrow\mathscr{I}$
maps every real-valued probability $x\in\left[0,1\right]$ to a set
of intervals $f\left(x\right)=\left[\ell,r\right]$ containing $x$,
where $\left[0,1\right]$ denotes the set of real-valued probabilities
which should not be considered as the interval-valued probability
$\unknown$. We also need a notion of norm to quantify the
distance between (pure or mixed) states. The norm of a pure state
$\rho=\proj{\psi}$ is defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
In addition, given any Hermitian operator~$A$, we consider the usual
operator norm $\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
the norm is the maximum absolute value of its eigenvalue. Then, a
finite-precision extension of Gleason's theorem can be stated as follows.

\begin{thm}[Finite-Precision Extension of the Gleason Theorem]\label{thm:Finite-precision-Gleason}Let
$f:\left[0,1\right]\rightarrow\mathscr{I}$ be a {\color{blue}interval map}
and let $f\circ\muB_{\rho}$ be a QIVPM, where $\muB_{\rho}$ is the
probability measure induced by the Born rule for a given state~$\rho$.
Let $\alpha$ be the maximum length of intervals in $\mathscr{I}$.
If a state $\rho'$ is consistent with $f\circ\muB_{\rho}$ on all
events, i.e., $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
then the norm of their difference is bounded by $\alpha$, i.e., $\left\Vert \rho-\rho'\right\Vert \le\alpha$.\end{thm}
\end{newText}
\end{compare}

@@

To summarize, a conventional quantum probability measure has an
uncountable range $[0,1]$. A QIVPM constructed by blurring \yutsung{Should we
change this ``blurring''?} such a
conventional quantum probability measure must also have an uncountable
range of intervals. Of course, any particular QIVPM, or any particular
experiment, will use a fixed collection of intervals appropriate for
the resources and precision of the particular experiment.
